---
# IMPORTANT: Change settings here, but DO NOT change the spacing.
# Remove comments and add values where applicable.
# The descriptions below should be self-explanatory

title: "Investigation of high South African Bond Yields"
#subtitle: "This will appear as Right Header"

documentclass: "elsarticle"

# --------- Thesis title (Optional - set to FALSE by default).
# You can move the details below around as you please.
Thesis_FP: FALSE
# Entry1: "An unbelievable study with a title spanning multiple lines."
# Entry2: "\\textbf{Nico Katzke}" # textbf for bold
# Entry3: "A thesis submitted toward the degree of Doctor of Philosophy"
# Uni_Logo: Tex/Logo.png # Place a logo in the indicated location (from your root, e.g. defaults to ~/Tex/Logo.png) and uncomment this line. Leave uncommented for no image
# Logo_width: 0.3 # If using a logo - use this to set width (size) of image
# Entry4: "Under the supervision of: \\vfill Prof. Joe Smith and Dr. Frank Smith"
# Entry5: "Stellenbosch University"
# Entry6: April 2020
# Entry7:
# Entry8:

# --------- Front Page
# Comment: ----- Follow this pattern for up to 5 authors
AddTitle: TRUE # Use FALSE when submitting to peer reviewed platform. This will remove author names.
Author1: "Wouter Bezuidenhout"  # First Author - note the thanks message displayed as an italic footnote of first page.
Ref1: "Stellenbosch University, South Africa" # First Author's Affiliation
Email1: "20088418\\@sun.ac.za" # First Author's Email address

# Author2: "John Smith"
# Ref2: "Some other Institution, Cape Town, South Africa"
# Email2: "John\\@gmail.com"
# CommonAffiliation_12: TRUE # If Author 1 and 2 have a common affiliation. Works with _13, _23, etc.

# Author3: "John Doe"
# Email3: "Joe\\@gmail.com"

CorrespAuthor_1: TRUE  # If corresponding author is author 3, e.g., use CorrespAuthor_3: TRUE

# Comment out below to remove both. JEL Codes only given if keywords also given.
keywords: "Bond Yields \\sep Emerging Markets" # Use \\sep to separate
JELCodes: "L250 \\sep L100"

# ----- Manage headers and footers:
#BottomLFooter: $Title$
#BottomCFooter:
#TopLHeader: \leftmark # Adds section name at topleft. Remove comment to add it.
BottomRFooter: "\\footnotesize Page \\thepage" # Add a '#' before this line to remove footer.
addtoprule: TRUE
addfootrule: TRUE               # Use if footers added. Add '#' to remove line.

# --------- page margins:
margin: 2.3 # Sides
bottom: 2 # bottom
top: 2.5 # Top
HardSet_layout: TRUE # Hard-set the spacing of words in your document. This will stop LaTeX squashing text to fit on pages, e.g.
# This is done by hard-setting the spacing dimensions. Set to FALSE if you want LaTeX to optimize this for your paper.

# --------- Line numbers
linenumbers: FALSE # Used when submitting to journal

# ---------- References settings:
# You can download cls format here: https://www.zotero.org/ - simply search for your institution. You can also edit and save cls formats here: https://editor.citationstyles.org/about/
# Hit download, store it in Tex/ folder, and change reference below - easy.
bibliography: Tex/ref.bib       # Do not edit: Keep this naming convention and location.
csl: Tex/harvard-stellenbosch-university.csl # referencing format used.
# By default, the bibliography only displays the cited references. If you want to change this, you can comment out one of the following:
#nocite: '@*' # Add all items in bibliography, whether cited or not
# nocite: |  # add specific references that aren't cited
#  @grinold2000
#  @Someoneelse2010

# ---------- General:
RemovePreprintSubmittedTo: TRUE  # Removes the 'preprint submitted to...' at bottom of titlepage
Journal: "Journal of Finance"   # Journal that the paper will be submitting to, if RemovePreprintSubmittedTo is set to TRUE.
toc: FALSE                       # Add a table of contents
numbersections: TRUE             # Should sections (and thus figures and tables) be numbered?
fontsize: 11pt                  # Set fontsize
linestretch: 1.2                # Set distance between lines.
link-citations: TRUE            # This creates dynamic links to the papers in reference list.

### Adding additional latex packages:
# header-includes:
#    - \usepackage{colortbl} # Add additional packages here.

output:
  pdf_document:
    keep_tex: TRUE
    template: Tex/TexDefault.txt
    fig_width: 3.5 # Adjust default figure sizes. This can also be done in the chunks of the text.
    fig_height: 3.5
abstract: |
  South African yield spreads in mid to longer dated bonds are currently relatively high to the historical experience. 
---

<!-- First: Set your default preferences for chunk options: -->

<!-- If you want a chunk's code to be printed, set echo = TRUE. message = FALSE stops R printing ugly package loading details in your final paper too. I also suggest setting warning = FALSE and checking for warnings in R, else you might find ugly warnings in your paper. -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.width = 6, fig.height = 5, fig.pos="H", fig.pos = 'H')
# Note: Include = FALSE implies the code is executed, but not printed in your pdf.
# warning and message = FALSE implies ugly messages and warnings are removed from your pdf.
# These should be picked up when you execute the command chunks (code sections below) in your rmd, not printed in your paper!

library(tidyverse)
library(xts)
library(rmsfuns)
library(fmxdat)
library(tbl2xts)
library(lubridate)

```


<!-- ############################## -->
<!-- # Start Writing here: -->
<!-- ############################## -->

# Introduction 

It was pointed out by economists that current yield spreads in local mid to longer dated bond yields have been historically high. Yield spreads between bonds of different maturities reflect how investors view the economic conditions. For example, widening spreads indicate largely different views between short to long term, and illustrates a view of better long term conditions than present. 

The price of a bond moves inversely to its yield, therefore a riskier bond is cheaper in price but greater in yield. Yield spreads are an indication of market sentiment. Relevant to the South African case, if investors are risk averse, they favor safer bonds, and therefore the spread between developed market bonds and emerging market bonds widen. 

Furthermore, yield differentials between bonds of the same class but with different maturities are also helpful indicators of sentiment. Longer-dated debt usually is considered riskier than short-dated debt, owing to the maturity risk premium. So, long-term debt typically would have a positive yield spread relative to short-term debt. But, at turning points in the market cycle, when a recession is priced in and fear of the immediate future prevails, investors shun shorter-term debt in favor of longer-term debt, thus pushing up short-term yields and dampening long-term yields.

In this paper, I review these avenues to investigate the higher-than-usual yield spreads in South Africa. 

# Historical Overview 

The first plot below shows the historical view of yields from 1999 to 2021, whilst the second plot shows the historical view of yield spreads over the same period in South Africa. The bond yields included are the 3-month, 2-year and 10-year government bonds. 

The first plot shows an increasing spike in the 10-year yield from the end of 2019 or start of 2020. Increasing yields during the period of Covid-19 is expected for a few reasons. There is a global risk of sentiment, so investors move out of EME bonds, this reduces bond prices and increase the yields. Secondly, the bonds are more risky as treasury embarked implemented fiscal stimulus amid already existing high debt levels. The 3-month bond and 2 year bond yields decrease over the same period as the SARB implement interest rate cuts and for the first time, intervene along the yield curve with some sense of quantatitive easing. 

The second plot shows that current yield spreads are higher than the historical yield spreads. The yield spreads spike at the end of 2019 or start of 2020, and stabilise at high levels. The spread between the 10-year and 3 month bonds are the greatest, followed by the difference between the 10 year and 2 year, and then the 2 year and 3 month. As shown in the first plot, the 10-year yield spiked upwards, while the 3-month and 2-year fall dramatically downwards. This causes the initial spike upwards that one observes in the yield spreads where the 10-year bond is involved, while the spread between the 3-month and 2-year is more tame. In order to understand this better, I review want to get an idea of the international investor sentiment. 


```{r}
SA_bonds <- read_rds("data/SA_Bonds.rds")
SA_bond_calc_1 <- SA_bonds %>% arrange(date) %>% gather(rate_type, rate, -date) %>%
ggplot() +
geom_line(aes(date, rate, colour = rate_type), size = 1.1, alpha = 0.8) +
fmxdat::theme_fmx() +
labs(x = "Date", y = "Yields", title = "Yields in South Africa", subtitle = "Historical view of yield spreads in South Africa, 1999 to 2021") +
fmx_cols()

finplot(SA_bond_calc_1, x.date.dist = "3 year", x.date.type = "%Y", x.vert = T, y.pct = F, y.pct_acc = 1)


```


```{r}

SA_bond_calc <- SA_bonds %>% arrange(date) %>% mutate(YS_10Yr_3M = ZA_10Yr - SA_3M) %>%
mutate(YS_10Yr_2Yr = ZA_10Yr - ZA_2Yr) %>%
mutate(YS_2yr_3M = ZA_2Yr - SA_3M) %>%
select(-SA_3M, -ZA_10Yr, -ZA_2Yr) %>%
gather(rate_type, rate, -date)

g <- SA_bond_calc %>%  ggplot() +
geom_line(aes(date, rate, colour = rate_type), size = 1.1, alpha = 0.8) +
fmxdat::theme_fmx() +
labs(x = "Date", y = "Yield Spreads", title = "Yield Spreads in South Africa", subtitle = "Historical view of yield spreads in South Africa, 1999 to 2021", caption = "Spreads are calculated as longer period less shorter period.") +
fmx_cols()

finplot(g, x.date.dist = "3 year", x.date.type = "%Y", x.vert = T, y.pct = F, y.pct_acc = 1)

```


# International Sentiment 

Firstly, I plot the USD/ZAR exchange rate to obtain an idea of volatility in carry trade and capital outflows. This figure shows a large spike in the exchange rate to levels above R18 per dollar. This is an indication of substantial capital outlflow out of South Africa. This is anecdotal evidence to expect that demand for SA bonds decreased largely by international investors. I argue that the reason we do not observe this as in the 3-month or 2-year is because of the substantial SARB intervention. 

```{r}
usdzar <- read_rds("data/usdzar.rds") %>%  filter(date > ymd(20180101))
names(usdzar)[3] <- 'USDZAR'
p <- usdzar %>% select (-Name) %>%
ggplot() +
geom_line(aes(date, USDZAR), color = "steelblue", size = 1.1, alpha = 0.8)
p
```

To confirm this finding, I compare the yield spread between the SA and US 10-year bonds, and the yield spread between the SA and US 2-year bonds. Fedderke (2021) argue that higher South African economic growth, lower inflation, public and private debt, as well as Rand-Dollar appreciation are all associated with a statistically significantly lower South African - United States yield spread. In addition to that, I would add global risk sentiment as contributing to yield spread evolution between SA and the US. 

The figure shows a clear spike in the yield spread between the US and SA. The spread is calculated as the SA yield less the US yield. This confirms that global risk sentiment, movement out of emerging market bonds and into safer developed markets bond is a contributing factor to higher yields in South Africa. In addition to this, I have omitted an discussion of an important factor being inflation. Investors are concerned about real yields, and therefore, if inflation expectations increase, then investors demand higher nominal yields to ensure constant real yields. Initially, there was a fear of deflation in the US, but this fear became less apparent as fiscal and monetary policy makers adopted great expansionary policy. 


```{r}
bonds_10yr <- read_rds("data/bonds_10y.rds")
bonds_2yr <- read_rds("data/bonds_2y.rds")

US_10yr <- bonds_10yr %>% group_by(Name) %>% filter(Name == "US_10Yr") %>% ungroup() %>% select(-Name)
names(US_10yr)[2] <- 'US_10Yr'
US_2yr <- bonds_2yr %>% group_by(Name) %>% filter(Name == "US_2yr") %>% ungroup() %>% select(-Name)
names(US_2yr)[2] <- 'US_2Yr'

# I want to compare SA yield to USA yield for maturities available. 

US_Yields <- left_join(US_10yr, US_2yr, by = "date")
SA_Yields <- read_rds("data/SA_Bonds.rds") %>% select(-SA_3M)
Combined_Yields <- left_join(US_Yields, SA_Yields, by = "date") %>% filter(date > ymd(20000101))

Combined_Spreads <- Combined_Yields %>% mutate(ZA_US_10Yr = ZA_10Yr - US_10Yr) %>%
mutate(ZA_US_2Yr = ZA_2Yr - US_2Yr) %>%  select(date, ZA_US_10Yr, ZA_US_2Yr) %>%
gather(Spread_Type, Spread, -date) %>% 
ggplot() +
geom_line(aes(date, Spread, colour = Spread_Type), size = 1.1, alpha = 0.8) +
fmxdat::theme_fmx() +
labs(x = "date", y = "Yield Spread", title = "Yield Spreads between South African and US Bonds", subtitle = "Historical view of yield spreads between South Africa and the USA, 2000 to 2021", caption = "Spreads are calculated as SA yield less USA yield.") +
fmx_cols()

finplot(Combined_Spreads, x.date.dist = "3 year", x.date.type = "%Y", x.vert = T, y.pct = F, y.pct_acc = 1)

```

# Conclusion

My main line of argument is that capital outflows into safer developed market bonds caused the demand for South African bonds to decrease, and therefore their yields to increase. I argue that the substantial action by the SARB, especially its adoption of a mild form quantative easing for the first time, kept the 3-month and 2-year yields low. However, the 10-year yield increased, and in turn its spread with bonds of shorter maturity. Higher yields for expected for longer-term bonds, but I do agree with the initial hypothesis that the yield spreads are greater than historical norms. 

<!-- Make title of bibliography here: -->
<!-- \newpage -->

\newpage

# References {-}

Fedderke, J.W., 2021. The South African–United States sovereign bond spread and its association with macroeconomic fundamentals. South African Journal of Economics.




