---
# IMPORTANT: Change settings here, but DO NOT change the spacing.
# Remove comments and add values where applicable.
# The descriptions below should be self-explanatory

title: "Comparison and Contrast of the ALSI (J200) and SWIX (J400)"
#subtitle: "This will appear as Right Header"

documentclass: "elsarticle"

# --------- Thesis title (Optional - set to FALSE by default).
# You can move the details below around as you please.
Thesis_FP: FALSE
# Entry1: "An unbelievable study with a title spanning multiple lines."
# Entry2: "\\textbf{Nico Katzke}" # textbf for bold
# Entry3: "A thesis submitted toward the degree of Doctor of Philosophy"
# Uni_Logo: Tex/Logo.png # Place a logo in the indicated location (from your root, e.g. defaults to ~/Tex/Logo.png) and uncomment this line. Leave uncommented for no image
# Logo_width: 0.3 # If using a logo - use this to set width (size) of image
# Entry4: "Under the supervision of: \\vfill Prof. Joe Smith and Dr. Frank Smith"
# Entry5: "Stellenbosch University"
# Entry6: April 2020
# Entry7:
# Entry8:

# --------- Front Page
# Comment: ----- Follow this pattern for up to 5 authors
AddTitle: TRUE # Use FALSE when submitting to peer reviewed platform. This will remove author names.
Author1: "Wouter Bezuidenhout"  # First Author - note the thanks message displayed as an italic footnote of first page.
Ref1: "Stellenbosch, South Africa" # First Author's Affiliation
Email1: "20088418\\@sun.ac.za" # First Author's Email address

# Author2: "John Smith"
# Ref2: "Some other Institution, Cape Town, South Africa"
# Email2: "John\\@gmail.com"
# CommonAffiliation_12: TRUE # If Author 1 and 2 have a common affiliation. Works with _13, _23, etc.

# Author3: "John Doe"
# Email3: "Joe\\@gmail.com"

CorrespAuthor_1: TRUE  # If corresponding author is author 3, e.g., use CorrespAuthor_3: TRUE

# Comment out below to remove both. JEL Codes only given if keywords also given.
keywords: "Index \\sep JSE" # Use \\sep to separate
JELCodes: "L250 \\sep L100"

# ----- Manage headers and footers:
#BottomLFooter: $Title$
#BottomCFooter:
#TopLHeader: \leftmark # Adds section name at topleft. Remove comment to add it.
BottomRFooter: "\\footnotesize Page \\thepage" # Add a '#' before this line to remove footer.
addtoprule: TRUE
addfootrule: TRUE               # Use if footers added. Add '#' to remove line.

# --------- page margins:
margin: 2.3 # Sides
bottom: 2 # bottom
top: 2.5 # Top
HardSet_layout: TRUE # Hard-set the spacing of words in your document. This will stop LaTeX squashing text to fit on pages, e.g.
# This is done by hard-setting the spacing dimensions. Set to FALSE if you want LaTeX to optimize this for your paper.

# --------- Line numbers
linenumbers: FALSE # Used when submitting to journal

# ---------- References settings:
# You can download cls format here: https://www.zotero.org/ - simply search for your institution. You can also edit and save cls formats here: https://editor.citationstyles.org/about/
# Hit download, store it in Tex/ folder, and change reference below - easy.
bibliography: Tex/ref.bib       # Do not edit: Keep this naming convention and location.
csl: Tex/harvard-stellenbosch-university.csl # referencing format used.
# By default, the bibliography only displays the cited references. If you want to change this, you can comment out one of the following:
#nocite: '@*' # Add all items in bibliography, whether cited or not
# nocite: |  # add specific references that aren't cited
#  @grinold2000
#  @Someoneelse2010

# ---------- General:
RemovePreprintSubmittedTo: TRUE  # Removes the 'preprint submitted to...' at bottom of titlepage
Journal: "Journal of Finance"   # Journal that the paper will be submitting to, if RemovePreprintSubmittedTo is set to TRUE.
toc: FALSE                       # Add a table of contents
numbersections: TRUE             # Should sections (and thus figures and tables) be numbered?
fontsize: 11pt                  # Set fontsize
linestretch: 1.2                # Set distance between lines.
link-citations: TRUE            # This creates dynamic links to the papers in reference list.

### Adding additional latex packages:
# header-includes:
#    - \usepackage{colortbl} # Add additional packages here.

output:
  pdf_document:
    keep_tex: TRUE
    template: Tex/TexDefault.txt
    fig_width: 3.5 # Adjust default figure sizes. This can also be done in the chunks of the text.
    fig_height: 3.5
abstract: |
  The report provides a comparison of the ALSI and SWIX Top 40 indexes. 
---

<!-- First: Set your default preferences for chunk options: -->

<!-- If you want a chunk's code to be printed, set echo = TRUE. message = FALSE stops R printing ugly package loading details in your final paper too. I also suggest setting warning = FALSE and checking for warnings in R, else you might find ugly warnings in your paper. -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.width = 6, fig.height = 5, fig.pos="H", fig.pos = 'H')

library(tidyverse)


```


<!-- ############################## -->
<!-- # Start Writing here: -->
<!-- ############################## -->

# Introduction 

This reports compares the ALSI (J200) and the SWIX (J400) top 40 Indexes. Both indexes operate consist of the 40 most investable companies in the ALSI universe, and both operate an free float market cap weighting methodlogy. The key distinction between the two indices is that the SWIX differs by using the STRATE
(South Africa’s Central Securities Depository) share register to reduce constituent weights of foreign or inward-listed companies.  The SWIX indices are therefore intended to be an indication of the shares held locally. 

The constituents of the SWIX 40 will be selected using SWIX-adjusted market capitalisation, and therefore the constituents of the SWIX 40 may differ from those of the ALSI Top 40 index. Investors have raised concerns about the high level of concentration of individual securities within the indices. This raises a discussion of capping individual securities to allow investors greater exposure to the rest of the market. This report investigates the differences between the two indexes. 


# Overall Comparison between SWIX and ALSI Top 40

The two Top 40 indexes are largely similar with small differences. The figure below compares their historical performance, and it is clear to see how synchronous the two indexes are. 

```{r}
# Data Wrangling 

library(pacman)
library(rmsfuns)
library(fmxdat)
p_load("tidyr","tbl2xts", "devtools", "lubridate","readr","PerformanceAnalytics", "tidyverse")
#Data
T40 <- read_rds("data/T40.rds")
RebDays <- read_rds("data/Rebalance_days.rds")

# Compare overall performance of SWIX and ALSI

R_xts <- T40 %>% select(-Short.Name, -Sector, -Index_Name, -J400, -J200) %>%
tbl_xts(., cols_to_xts = Return, spread_by = Tickers)

J200_W_xts <- T40 %>% select(-Short.Name, -Sector, -Index_Name, -J400, -Return) %>%
tbl_xts(., cols_to_xts = J200, spread_by = Tickers)

J400_W_xts <- T40 %>% select(-Short.Name, -Sector, -Index_Name, -J200, -Return) %>%
tbl_xts(., cols_to_xts = J400, spread_by = Tickers)

# Set all NA returns to zero:
R_xts[is.na(R_xts)] <- 0
J400_W_xts[is.na(J400_W_xts)] <- 0
J200_W_xts[is.na(J200_W_xts)] <- 0

J200_Portfolio <- rmsfuns::Safe_Return.portfolio(R = R_xts, weights = J200_W_xts, geometric = TRUE)
J400_Portfolio <- rmsfuns::Safe_Return.portfolio(R = R_xts, weights = J400_W_xts, geometric = TRUE)

J200_Portf_Rets <- J200_Portfolio$portfolio.returns %>% xts_tbl() %>% mutate(J200_Cum_Ret = cumprod(1+portfolio.returns))

J400_Portf_Rets <- J400_Portfolio$portfolio.returns %>% xts_tbl() %>% mutate(J400_Cum_Ret = cumprod(1+portfolio.returns))

port_compare <- left_join(J200_Portf_Rets, J400_Portf_Rets, by = "date") %>% select(date, J200_Cum_Ret, J400_Cum_Ret) %>%
gather(Index, Return, -date) 

plot_port <- port_compare %>%  ggplot() + geom_line(aes(date, Return, color = Index)) + theme_bw() +
fmxdat::fmx_cols() +
fmxdat::theme_fmx(title.size = ggpts(50), CustomCaption = F) +
labs(x = "Date", y = "Cumulative Returns", title = "J200 and J400 Cumulative Returns over time",
subtitle = "A comparison of the cumulative returns of the ALSI Top 40 and the SWIX Top 40 over time")

plot_port



```

In order to investigate this a bit further, I present the following figure showing the historic performance of the indexes' allocation to large caps using the exact weights as used by the index. As to be expected from a top 40 index, the majority of allocation are to large caps. If one compares the above figure to the figure below, under close inspection,  one can see that the magnitude of the y-axis is large in the full version (above) relative to the large-cap version. Small caps, intuitively, do not feature in top 40 indexes. 

```{r}

#### Compare overall performance over Large Caps ##########

LC_R_xts <- T40 %>% select(-Short.Name, -Sector, -J400, -J200) %>% filter(Index_Name == "Large_Caps") %>%
tbl_xts(., cols_to_xts = Return, spread_by = Tickers)

LC_J200_W_xts <- T40 %>% select(-Short.Name, -Sector, -J400, -Return) %>% filter(Index_Name == "Large_Caps") %>%
tbl_xts(., cols_to_xts = J200, spread_by = Tickers) 

LC_J400_W_xts <- T40 %>% select(-Short.Name, -Sector, -Sector, -J200, -Return) %>% filter(Index_Name == "Large_Caps") %>%
tbl_xts(., cols_to_xts = J400, spread_by = Tickers)

# Set all NA returns to zero:
LC_R_xts[is.na(LC_R_xts)] <- 0
LC_J400_W_xts[is.na(LC_J400_W_xts)] <- 0
LC_J200_W_xts[is.na(LC_J200_W_xts)] <- 0

LC_J200_Portfolio <- rmsfuns::Safe_Return.portfolio(R = LC_R_xts, weights = LC_J200_W_xts, geometric = TRUE)
LC_J400_Portfolio <- rmsfuns::Safe_Return.portfolio(R = LC_R_xts, weights = LC_J400_W_xts, geometric = TRUE)

LC_J200_Portf_Rets <- LC_J200_Portfolio$portfolio.returns %>% xts_tbl() %>% mutate(LC_J200_Cum_Ret = cumprod(1+portfolio.returns))

LC_J400_Portf_Rets <- LC_J400_Portfolio$portfolio.returns %>% xts_tbl() %>% mutate(LC_J400_Cum_Ret = cumprod(1+portfolio.returns))

LC_port_compare <- left_join(LC_J200_Portf_Rets, LC_J400_Portf_Rets, by = "date") %>% select(date, LC_J200_Cum_Ret, LC_J400_Cum_Ret) %>%
    gather(Index, Return, -date) 

LC_plot_port <- LC_port_compare %>%  ggplot() + geom_line(aes(date, Return, color = Index)) + theme_bw() +
    fmxdat::fmx_cols() +
    fmxdat::theme_fmx(title.size = ggpts(50), CustomCaption = F) +
    labs(x = "Date", y = "Cumulative Returns", title = "Large Caps J200 and J400 Cumulative Returns over time",
         subtitle = "A comparison of the cumulative returns of the Large Caps in the ALSI Top 40 and the SWIX Top 40 over time")
LC_plot_port
# fmxdat::finplot(LC_port_compare, x.vert = T, x.date.type = "%Y", x.date.dist = "1 years")

```

I turn to the medium-caps in order to draw a more distinct comparison between the two top 40 indexes. An interesting pattern emerges here, in the previous two figures, the J200 outperforms the J400 in 2021 (cumulatively), but with the midcaps, the J400 outperforms the J200. This could either be because the J400 has greatest exposure to midcaps or has the same exposure, but includes better midcaps. 

```{r}

MC_R_xts <- T40 %>% select(-Short.Name, -Sector, -J400, -J200) %>% filter(Index_Name == "Mid_Caps") %>%
    tbl_xts(., cols_to_xts = Return, spread_by = Tickers)

MC_J200_W_xts <- T40 %>% select(-Short.Name, -Sector, -J400, -Return) %>% filter(Index_Name == "Mid_Caps") %>%
    tbl_xts(., cols_to_xts = J200, spread_by = Tickers) 

MC_J400_W_xts <- T40 %>% select(-Short.Name, -Sector, -Sector, -J200, -Return) %>% filter(Index_Name == "Mid_Caps") %>%
    tbl_xts(., cols_to_xts = J400, spread_by = Tickers)

# Set all NA returns to zero:
MC_R_xts[is.na(MC_R_xts)] <- 0
MC_J400_W_xts[is.na(MC_J400_W_xts)] <- 0
MC_J200_W_xts[is.na(MC_J200_W_xts)] <- 0

MC_J200_Portfolio <- rmsfuns::Safe_Return.portfolio(R = MC_R_xts, weights = MC_J200_W_xts, geometric = TRUE)
MC_J400_Portfolio <- rmsfuns::Safe_Return.portfolio(R = MC_R_xts, weights = MC_J400_W_xts, geometric = TRUE)

MC_J200_Portf_Rets <- MC_J200_Portfolio$portfolio.returns %>% xts_tbl() %>% mutate(MC_J200_Cum_Ret = cumprod(1+portfolio.returns))

MC_J400_Portf_Rets <- MC_J400_Portfolio$portfolio.returns %>% xts_tbl() %>% mutate(MC_J400_Cum_Ret = cumprod(1+portfolio.returns))

MC_port_compare <- left_join(MC_J200_Portf_Rets, MC_J400_Portf_Rets, by = "date") %>% select(date, MC_J200_Cum_Ret, MC_J400_Cum_Ret) %>%
gather(Index, Return, -date) 

MC_plot_port <- MC_port_compare %>%  ggplot() + geom_line(aes(date, Return, color = Index)) + theme_bw() +
    fmxdat::fmx_cols() +
    fmxdat::theme_fmx(title.size = ggpts(50), CustomCaption = F) +
    labs(x = "Date", y = "Cumulative Returns", title = "Mid Caps J200 and J400 Cumulative Returns over time",
         subtitle = "A comparison of the cumulative returns of the Mid Caps in the ALSI Top 40 and the SWIX Top 40 over time")
MC_plot_port
# fmxdat::finplot(MC_port_compare, x.vert = T, x.date.type = "%Y", x.date.dist = "1 years")

```


# Volatility 

I present a simple calculation about volatility that each indix experiences in times where the USD/ZAR has high or low volatility. I measure the standard deviation of the USD/ZAR, classify periods within above the 80th percentile of standard deviation as high volatility, and below the 20th percentile as low volatility. The results are as follows. 


```{r}
library(kableExtra)
# Volatile USDZAR 
zar <- read_rds("data/usdzar.rds")

ZARSD <- zar %>% filter(date > ymd(20080102)) %>%
mutate(YearMonth = format(date,"%Y%B")) %>% group_by(YearMonth) %>% summarise(SD = sd(Price)*sqrt(52)) %>%
mutate(TopQtile = quantile(SD, 0.8), BotQtile = quantile(SD,0.2))

Hi_Vol <- ZARSD %>% filter(SD >TopQtile) %>% pull(YearMonth)
Low_Vol <- ZARSD %>% filter(SD < BotQtile) %>% pull (YearMonth)

## Who has better vol? 

overall_SD <- port_compare %>% mutate(YearMonth = format(date,"%Y%B")) %>% group_by(Index) %>%
summarise(Full_SD = sd(Return)*sqrt(252))

high_vol_SD <- port_compare %>% mutate(YearMonth = format(date,"%Y%B")) %>%
filter(YearMonth %in% Hi_Vol)%>% group_by(Index) %>%
summarise(Highvol_SD = sd(Return)*sqrt(252))
    
Low_vol_SD <- port_compare %>% mutate(YearMonth = format(date,"%Y%B")) %>%
filter(YearMonth %in% Low_Vol)%>% group_by(Index) %>%
summarise(Lowvol_SD = sd(Return)*sqrt(252))


vol_compare_demo <- left_join(overall_SD, high_vol_SD, by = "Index")

vol_compare <- left_join(vol_compare_demo, Low_vol_SD, by = "Index") %>%
mutate(high_SD_ratio = Highvol_SD/Full_SD, low_SD_Ratio = Lowvol_SD/Full_SD) %>% kable()
vol_compare


```

The calculation shows that the standard deviation (SD) of the J200 is slightly less than the J400 in the full sample period. Furthermore, the SD of the J200 is lower than that of the J400 in high and low volatility times. However, if one compares the ratios, high volatility SD divided by full SD for example, one sees that the J200 has a greater ratio than the J400 in high volatility times and a lower ratio in low volatility times. In a sense, this means SD can change faster but within a smaller bound. 

# Capping the Indexes 



# Conclusion

The two indexes are very similar, however it is known that the SWIX focuses more on holding securities that mimic local holdings as opposed to foreign holdings. 

<!-- Make title of bibliography here: -->
<!-- \newpage -->

\newpage

# References {-}

<div id="refs"></div>




